---
title: "Vignette - FarsDataAnalysisPD"
author: "Paul Draper"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FarsDataAnalysisPD}
  %\VignetteEngine{knitr::rmarkdown}  
  %\VignetteEncoding{UTF-8}
---

```{r setenv, echo=FALSE, message=FALSE}
library(FarsDataAnalysisPD)
library(magrittr)
```

This vignette provides a brief introduction to using the `FarsDataAnalysisPD` package. This package provides tools to read and analyze the National Highway Traffic Safety Administration's (NHTSA) Fatality Analysis Reporting System (FARS) data. This package contains the following functions:

* make_filename
* fars_read
* fars_read_years
* fars_summarize_years
* fars_map_state

## Raw Data

This package includes FARS data from three years: 2013, 2014, 2015. A snapshot of the 2013 data is as follows:

```{r 2013_RawData, echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv(system.file("extdata", "accident_2013.csv.bz2", package = "FarsDataAnalysisPD")) %>% 
  head(10)
```

In the next sections, we will see how to read in and process this data.

## Read in FARS data

FARS data can either be read in for a single year or multiple years. To read in a single year of data, first utilize the `make_filename` function to create a filename for the given year of interest. Then, use `fars_read` to read in the file.

```{r echo=TRUE}
setwd(file.path("..", "inst", "extdata"))
FARS_data_2013_file <- make_filename(2013)
FARS_data_2013 <- fars_read(FARS_data_2013_file)
```

To read multiple years of data, use the `fars_read_years` function:

```{r echo=TRUE, message=FALSE, warning=FALSE}
setwd(file.path("..", "inst", "extdata"))
FARS_data_multiple <- fars_read_years(2013:2015)
```

This creates a list containing the individual data files. Of course, this function could be used to read a single year of data as well.

## Summarizing FARS data

To summarize the number of accidents by month for a set of years, use the `fars_summarize_years` function. For example, to summarize the accidents for years 2013 - 2014, run

```{r}
setwd(file.path("..", "inst", "extdata"))
fars_summarize_years(2013:2014)
```

## Plot Accident Fatalities by State

The `fars_map_state` function takes a state ID number and year, and produces a map of accident fatalities for the given state. 

```{r}
setwd(file.path("..", "inst", "extdata"))
fars_map_state(1, 2013)
```

